---
title: "Replication of 'Does Seeing Faces of Young
Black Boys Facilitate the Identification of
Threatening Stimuli?' by Todd, Thiem, & Neel (2016, Psychological Science)"
author: "Cinoo Lee (cinoolee@stanford.edu)"
date: "2017.03"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: false
---


##Introduction

We will try to replicate the following result from the original study:

"Pervasive stereotypes linking Black men with violence and criminality can lead to implicit cognitive biases, including the misidentification of harmless objects as weapons. [..] We investigated whether these biases extend even to young Black boys (5-year-olds). Results [revealed] that participants had less difficulty (i.e., faster response times, fewer errors) identifying threatening stimuli and more difficulty identifying nonthreatening stimuli after seeing Black faces than after seeing White faces." (Todd et al., 2016)

##Methods

"White participants completed sequential priming tasks in which they categorized threatening and nonthreatening objects (toys and guns) and words after brief presentations of Black and White children's faces."

###Power Analysis

Original $\sf{η_{p}}$^2^ = .22 which translates into an effect size of 0.53. Power analysis shows that I need 16 participants for 80% power, 20 for 90%, and 24 for 95% power. This effect size is for the 2 (race of prime: Black, White) × 2 (target object: gun, toy) repeated measures analysis of variance (ANOVA).

###Planned Sample

Other - Planned sample size is 37. 99% power yields n = 32 and we added 5, which is 15% of 32 to ensure we will still have enough participants after the exclusion criteria. We will exclude non-White participants in the analysis.

###Materials

"The primes were 12 photos of boys (6 Black, 6 White) taken from the Child Affective Facial Expression set (LoBue & Thrasher, 2015). We selected these photos using the following criteria: The faces had to be easily categorized by race, to have a neutral expression, to have no idiosyncrasies (e.g., facial scars), and to be similar in actual age (mean age for Black faces = 4.98 years; mean age for White faces = 5.01 years; p > .250). Each photo was cropped so that it included only the head and was standardized in size. The target objects were 6 gun images taken from Payne (2001) and 6 toy images (e.g., a rattle) taken from online sources. The toy images were converted to gray scale and sized to match the gun images."

###Procedure	

Instead of beginning with a blank screen (500 ms) like in the original article, we began with a screen that instructed participants to "Press the spacebar to continue" after each trial. The rest of the procedures were the same.
"Each trial sequence began [..] followed by a face prime (200 ms), then a target object (200 ms), and finally a pattern mask (which remained on screen until participants responded). If participants did not respond within 500 ms, a message (“Please respond faster!”) appeared for 1 s. Each of the 12 face primes was paired once with each of the 12 target objects, which resulted in 144 randomly ordered experimental trials. Eight practice trials preceded the experimental trials."

###Analysis Plan

"We decided a priori to exclude data from participants with below-chance accuracy (i.e., errors on > 50% of trials) on the weapon identification task."

"Before analysis, we excluded errors and RTs less than 100 ms. RTs exceeding
the 500-ms deadline were excluded from analyses, and responses on those trials were treated as errors. We then subjected the remaining RTs to a log transformation (Payne, 2001); however, for interpretive ease, we report raw RTs.
A 2 (race of prime: Black, White) × 2 (target object: gun, toy) repeated measures analysis of variance (ANOVA) revealed a significant interaction." The key test will be the interaction effect of this 2 X 2 ANOVA.

We will only analyze data from people who indicated their race as White.

###Differences from Original Study

Difference in procedure is noted above. Instead of a lab study, this study will be conducted online. These differences are not anticipated to make a difference in the effect.

### Methods Addendum (Post Data Collection)


#### Actual Sample
  Sample size: 37
  Demographics: 
  Data exclusions:

#### Differences from pre-data collection methods plan
  None.

##Results


### Data preparation

```{r include = F}
####Load Relevant Libraries and Functions
#install.packages("ez")
library(tidyverse)
library(forcats)
library(rjson)
library(ez)
library(forcats)

####Import data
raw.data.path <- "~/Documents/Classes/Winter2017/PSYC254/Replication/Cosub/Todd2016/mturk/production-results/"

files <- dir(raw.data.path,pattern="*.json")
d.raw = NA
d.raw <- data.frame()


for (f in files) {
  jf <- paste0(raw.data.path,f)
  jd <- fromJSON(file=jf)
  

  id <- data.frame(workerid = jd$WorkerId, 
                   religAff = jd$answers$demographicsData[[1]]$ReligiousAffiliation,
                   religAtt = jd$answers$demographicsData[[1]]$ReligiousAttendance,
                   ses = jd$answers$demographicsData[[1]]$SES,
                   age = jd$answers$demographicsData[[1]]$Age,
#                   political = jd$answers$demographicsData[[1]]$Political,
                   comments = jd$answers$demographicsData[[1]]$Comments,
                   Edu = jd$answers$demographicsData[[1]]$Education,
                   Citizen = jd$answers$demographicsData[[1]]$Citizen,
                   Gender = jd$answers$demographicsData[[1]]$Male,
                   Ethnicity = jd$answers$demographicsData[[1]]$Ethnicity,
                   trial_num = 1:144,
                   rt = NA,
                   accuracy = NA,
                   raceStim = NA,
                   toolType= NA,
                   race = NA,
                   toolStim = NA,
                   responded = NA)

for(i in 1:144) {
    id$rt[id$trial_num == i] <- jd$answers$data[[i]]$rt
    id$raceStim[id$trial_num == i] <- jd$answers$data[[i]]$raceStim
    id$toolType[id$trial_num == i] <- jd$answers$data[[i]]$tool
    id$race[id$trial_num == i] <- jd$answers$data[[i]]$race
    id$toolStim[id$trial_num == i] <- jd$answers$data[[i]]$toolStim
    id$responded[id$trial_num == i] <- jd$answers$data[[i]]$responded
    id$accuracy[id$trial_num == i] <- jd$answers$data[[i]]$accuracy
  }
  
    d.raw <- bind_rows(d.raw, id)
    
}


summary(d.raw)

d <- d.raw %>%
  select(-race) %>%
  mutate(targetRace = substr(d.raw$raceStim,0,1))

d$rt[d$rt == 0] <- NA

d$toolType = factor(d$toolType) 
d$targetRace = factor(d$targetRace)
d$Edu = factor(d$Edu) 
d$Citizen = factor(d$Citizen)
d$Gender = factor(d$Gender)
d$Ethnicity = factor(d$Ethnicity)

###Factoring and re-labeling

d$toolType = fct_recode(d$toolType,
                      "Toy" = "t",
                        "Gun" = "g")

d$targetRace = fct_recode(d$targetRace,
                      "Black" = "b",
                        "White" = "w")
d$Edu = fct_recode(d$Edu,
                   "Some_hs" = "1",
                   "Completed_hs" = "2",
                   "Some_college" = "3",
                   "Bachelors_degree" = "4",
                   "Masters_Professional_Doctoral" = "5")

d$Citizen = fct_recode(d$Citizen,
                   "American Citizen" = "1",
                   "Permanent Resident of the United States" = "2",
                   "Foreign Citizen" = "3")

d$Gender = fct_recode(d$Gender,
                   "Male" = "1",
                   "Female" = "0")
                   
d$Ethnicity = fct_recode(d$Ethnicity,
                    "Caucasian/White" = "1",
                    "Hispanic/Latino/Latina" = "2",
                    "African American/Black" = "3",
                    "Hispanic/Latino/Latina" = "4",
                    "East Asian" = "5",
                    "Middle Eastern" = "6",
                    "Pacific Islander" = "7",
                    "Native American" = "8",
                    "Mixed" = "9")

# Number of participants
length(unique(d$workerid))
```

```{r}
###Exclusion 
#Exclude non-White participants
d.white <- d %>%
  filter(Ethnicity == "Caucasian/White")

length(unique(d.white$workerid))

#assess accuracy of people among those who have responded to the question
assess_accuracy  <- d %>%
  filter(responded==1) %>%
  group_by(workerid) %>%
summarise(mean_accuracy = mean(accuracy))

#assess n of participants who are 50% accuracy or higher
assess_accuracy  <- d %>%
  filter(responded==1) %>%
  group_by(workerid) %>%
  summarise(mean_accuracy = mean(accuracy)) %>%
  filter(mean_accuracy > 0.499)
#this comes out to be 21 people for all ethnicity and 16 people for Whites.
#adjusting the filter to 40% or higher makes everyone pass the bar.

d.complete.all <- d %>%
   filter(!(is.na(rt)) & accuracy == 1 & rt > 100 &  rt <= 500) %>%
   mutate(log_rt = log(rt))

d.complete.white <- d.white %>%
  filter(!(is.na(rt)) & accuracy == 1 & rt > 100 &  rt <= 500) %>%
   mutate(log_rt = log(rt))
```

###Plotting Preparation
```{r}
sem <- function(x) {sd(x) / sqrt(length(x))}

ci95.norm <- function(x) {
  me <- sem(x)*qnorm(.975)
  c(mean(x)-me, mean(x)+me)
}

means.all <- d.complete.all %>%
  filter(accuracy == 1) %>% #only want correct trials
  filter(!is.na(rt)) %>%
  group_by(workerid, targetRace, toolType) %>%
  summarise(participant_mean = mean(rt)) %>%
  group_by(targetRace, toolType, add=FALSE) %>%
  summarise(m = mean(participant_mean), ci.lo=ci95.norm(participant_mean)[1], ci.hi = ci95.norm(participant_mean)[2])

means.white <- d.complete.white %>%
  filter(accuracy == 1) %>% #only want correct trials
  filter(!is.na(rt)) %>%
  group_by(workerid, targetRace, toolType) %>%
  summarise(participant_mean = mean(rt)) %>%
  group_by(targetRace, toolType, add=FALSE) %>%
  summarise(m = mean(participant_mean), ci.lo=ci95.norm(participant_mean)[1], ci.hi = ci95.norm(participant_mean)[2])

limits = aes(ymax=ci.hi, ymin=ci.lo)
```

### Confirmatory analysis

```{r}
ggplot(means.all, aes(x=targetRace, y=m, fill=toolType)) +
  geom_bar(stat="identity", position="dodge", width=.9) + 
  geom_errorbar(limits, position=position_dodge(width=.9), width=0.05) +
  theme_bw() +
  ggtitle("Weapon Identification Reaction Times--All participants") +
  labs(x = "Target race", y = "Mean reaction time (ms)")

ggplot(means.white, aes(x=targetRace, y=m, fill=toolType)) +
  geom_bar(stat="identity", position="dodge", width=.9) + 
  geom_errorbar(limits, position=position_dodge(width=.9), width=0.05) +
  theme_bw() +
  ggtitle("Weapon Identification Reaction Times--White participants") +
  labs(x = "Target race", y = "Mean reaction time (ms)")

means.aov <- d.complete.all %>%
  filter(accuracy == 1) %>% #only want correct trials
  filter(!is.na(log_rt)) %>%
  group_by(workerid, targetRace, toolType) %>%
  summarise(meanRT = mean(log_rt))

aov <- with(means.aov, aov(meanRT ~ targetRace * toolType + Error(workerid/(targetRace * toolType))))

# anova
summary(means.aov)
View(means.aov)

# calculate PARTIAL ETA SQUARED
```
summary(aov)

#aov <- ezANOVA(data = d.complete.all, dv = log_rt, wid = workerid, within = .(targetRace, toolType), type = 2, detailed = TRUE)

###Exploratory analyses

# lmer
#refer
jso_mss <- jso_final %>% 
  filter(correct == 1) %>%
  filter(cond != "ap") %>%
  mutate(item = substr(leftPic, 1, nchar(leftPic)-1)) %>%
  group_by(cond, subid, item) %>%
  summarise(rt = mean(rt), na.rm = TRUE)

jso_mss$cond <- as.factor(jso_mss$cond)
jso_mss$cond <- relevel(jso_mss$cond, "ds")
lmer <- lmer(rt ~ cond + (1 | subid) + (1 | item), data = jso_mss)
summary(lmer)

## Discussion


### Summary of Replication Attempt

Open the discussion section with a paragraph summarizing the primary result from the confirmatory analysis and the assessment of whether it replicated, partially replicated, or failed to replicate the original result.  

### Commentary

(a) insights from follow-up exploratory analysis

(b) assessment of the meaning of the replication (or not) - e.g., for a failure to replicate, are the differences between original and present study ones that definitely, plausibly, or are unlikely to have been moderators of the result, and 

(c) discussion of any objections or challenges raised by the current and original authors about the replication attempt.  None of these need to be long.

The original author commented on 1) The cover page of the study: "This task is investigating object recognition under distracting conditions. You will be asked to quickly identify objects will being distracted by other stimuli.", 2) Warning message: "It might be helpful for participants if you warn them that they'll see a message (an X) if they fail to respond quickly enough.", and 3) Deviation: Requiring participants to press the spacebar between trials is one of the only deviations from the original procedure. 